---
title: "Mini-Project 2"
author: "Yvonne Niyonzima, Duanyue Yun, Junzhou Liu"
date: "March 24, 2017"
output: html_document
---


#Our goal was to create a code that can inform us the amount of money contributed to a party (demorats, independents, and republicans) in the senatorial races during elections years (this data can be used to study whether or not there exists a correlation between the amount contributed and the number of seats won).

```{r,  include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(tidyr)

```


```{r, include=FALSE}
load("house_elections.rda")
load("candidates.rda")
load("committees.rda")
load("contributions.rda")
```


```{r}
joined <- candidates %>%
  full_join(contributions, by = "cand_id")
#we first join the  data of contributiions and candidates by the "candidate id"
##We chose only candidates who are seeking their state senate seats in the years between 2010 and 2016 using the verb "filter". We also use select to have only the variables that we need. 
Sjoined <- joined %>%
  filter(cand_office == "S"  ) %>%
  select(cand_id, cand_party_affiliation, cand_election_yr, cand_ici, cand_state, transaction_amt) ##%>%
  
##we want to replace all the na with the 0
Sjoined[is.na(Sjoined)] <- 0

#we only choose the twomajor parties (dems and republicans), but we also add independents. We only did these three because they received most contributions.
Sjoined_party <- filter(Sjoined,cand_party_affiliation == "REP" | cand_party_affiliation =="DEM" | cand_party_affiliation == "IND"  )


theyear2012 <- filter(Sjoined_party, cand_election_yr == "2012")

summary2012 <- theyear2012 %>%
  group_by(cand_party_affiliation) %>%
  summarize(amount = sum(transaction_amt))  #This codes will show us the amount each party received for the senatorial race in the year 2012.

  plot <- ggplot(data = summary2012, aes(x = cand_party_affiliation, y = amount, fill=cand_party_affiliation)) +
  geom_bar(stat = "identity", aes())+
    scale_fill_manual(values = c("blue", "green", "red"))

```

#The visual is a bar graph of the year 2012. As it can be seen, the republican received more money than both democrats and independents combined.

```{r}
plot
```

#The code can be modified also to analyze the correlation of the amount received in any political race. 